KidsAIR initial SUMs data work

### 2020-01-03 sums_data_cleaning.Rmd
Clean SUMs data

---
title: "KidsAIR initial PM data work"
author: "Ethan Walker"
date: "Started 23 Dec 2019, Updated 27 Dec 2019"
output: pdf_document
---

@@

# Add linked IDs to WMT SUMs data logs
```{r}
wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date = InstallDate,
         time = InstallTime) %>% 
  select(home_winter_id, ibid, date, time)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "WMT") %>% 
  select(home_winter_id, home, home_id)

wmt_sums_log <- wmt_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id, date, time, ibid) %>% 
  arrange(home, date)

write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")

@@

# Add linked IDs to NN SUMs data logs
```{r}
nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date = InstallDate,
         time = InstallTime) %>% 
  select(home_winter_id, ibid, date, time)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "NN") %>% 
  select(home_winter_id, home, home_id)

nn_sums_log <- nn_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id, date, time, ibid) %>% 
  arrange(home, date)

write_csv(nn_sums_log, "Output/nn_sums_data_log.csv")

@@

# Add linked IDs to AK SUMs data logs
```{r}
ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date = InstallDate,
         time = InstallTime) %>% 
  select(home_winter_id, ibid, date, time)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "AK") %>% 
  select(home_winter_id, home, home_id)

ak_sums_log <- ak_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, home, home_id, date, time, ibid) %>% 
  arrange(home, date)

write_csv(ak_sums_log, "Output/ak_sums_data_log.csv")

@@

# Bind all WMT SUMs data together
ak_sums_raw <- rbind(ak_sums_csv, ak_sums_csv_2, 
                     ak_sums_xlsx, ak_sums_xlsx2) %>% 
  select(-datetime_sums) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  filter(!is.na(date_sums)) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
```

# Code from PM cleaning
```{r}
##### Read in dusttrack .csv files (all areas) #####

### WMT ###
# List files
list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "WMT") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

### NN ###
# List files
list_csv_files <- list.files("Input/NN/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/NN/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

### AK ###
# List files
list_csv_files <- list.files("Input/AK/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/AK/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
ak_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "AK") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

##### Read in other dusttrack .csv files (WMT and NN) #####

### WMT ### 
# List files
list_other_files <- list.files("Input/WMT/dusttrack_other_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/dusttrack_other_files")
initial_data = tibble(file_list = list_other_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_pm_other <- initial_data %>% 
  mutate(trash1 = `Instrument Name`,
         trash2 = `DustTrak II`) %>% 
  select(home_winter_id, trash1, trash2) %>% 
  group_by(home_winter_id) %>% 
  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash2, "NA"),
         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash1, "NA"),
         trash1 = as.numeric(trash1),
         trash1 = as.character(trash1),
         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
  filter(!is.na(trash5)) %>% 
  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
  spread(trash4, trash3) %>% 
  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
  mutate(datetime = lead(datetime)) %>% 
  replace_with_na(replace = list(datetime = "NA NA")) %>% 
  mutate(datetime2 = first(datetime),
         datetime = mdy_hms(datetime2),
         trash1 = as.numeric(trash1)) %>% 
  filter(!is.na(trash1)) %>% 
  mutate(datetime2 = datetime + seconds(trash1),
         datetime2 = ymd_hms(datetime2)) %>% 
  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(pm = as.numeric(trash2),
         datetime = datetime2,
         area = "WMT") %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  ungroup()

### NN ### 
# List files
list_other_files <- list.files("Input/NN/dusttrack_other_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/NN/dusttrack_other_files")
initial_data = tibble(file_list = list_other_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_pm_other <- initial_data %>% 
  mutate(trash1 = `Instrument Name`,
         trash2 = `DustTrak II`) %>% 
  select(home_winter_id, trash1, trash2) %>% 
  filter(!is.na(trash1)) %>% 
  group_by(home_winter_id) %>% 
  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash1, "NA"),
         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash2, "NA"),
         trash1 = as.numeric(trash1),
         trash1 = as.character(trash1),
         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
  filter(!is.na(trash5)) %>% 
  spread(trash3, trash4) %>% 
  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
  mutate(datetime = lead(datetime)) %>% 
  replace_with_na(replace = list(datetime = "NA NA")) %>% 
  mutate(datetime2 = first(datetime),
         datetime = mdy_hms(datetime2),
         trash1 = as.numeric(trash1)) %>% 
  filter(!is.na(trash1)) %>% 
  mutate(datetime2 = datetime + seconds(trash1),
         datetime2 = ymd_hms(datetime2)) %>% 
  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(pm = as.numeric(trash2),
         datetime = datetime2,
         area = "NN") %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  ungroup() %>% 
  mutate(home_winter_id = if_else(home_winter_id == "220001" | home_winter_id == "220002",
                                  "22", home_winter_id),
         home_winter_id = if_else(home_winter_id == "550001" | home_winter_id == "550002",
                                  "55", home_winter_id),
         home_winter_id = if_else(home_winter_id == "810001" | home_winter_id == "810002",
                                  "81", home_winter_id))

##### Bind all files from above #####
kids_pm <- rbind(wmt_pm_csv, wmt_pm_txt, wmt_pm_other,
                nn_pm_csv, nn_pm_other,
                ak_pm_csv)

# Save pm dataset as RDS and CSV
write_rds(kids_pm, "Output/kids_pm_full_raw.rds")
#write_csv(kids_pm, "Output/kids_pm_full.csv")
```

>>>>

---
title: "KidsAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 23 Dec 2019, Updated 3 Jan 2020"
output: pdf_document
---

@@

# Add linked IDs to WMT SUMs data logs
```{r}
wmt_sums_data_log <- read_csv("Input/WMT/wmt_sums_data_log.csv") %>%
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "WMT") %>% 
  select(home_winter_id, winter_id, home, home_id)

wmt_sums_log <- wmt_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
  arrange(home, date_log) %>% 
  mutate(date_log = mdy(date_log))

write_rds(wmt_sums_log, "Output/wmt_sums_data_log.rds")
write_csv(wmt_sums_log, "Output/wmt_sums_data_log.csv")

@@

# Add linked IDs to NN SUMs data logs
```{r}
nn_sums_data_log <- read_xlsx("Input/NN/nn_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  select(home_winter_id, ibid, date_log, time_log)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "NN") %>% 
  select(home_winter_id, winter_id, home, home_id)

nn_sums_log <- nn_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
  arrange(home, date_log) %>% 
  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
  select(-trash)

write_rds(nn_sums_log, "Output/nn_sums_data_log.rds")
write_csv(nn_sums_log, "Output/nn_sums_data_log.csv")

@@

# Add linked IDs to AK SUMs data logs
```{r}
ak_sums_data_log <- read_xlsx("Input/AK/ak_sums_data_log.xlsx") %>% 
  mutate(home_winter_id = as.character(HomeWinterID),
         ibid = iButtonID,
         date_log = InstallDate,
         time_log = InstallTime) %>% 
  separate(time_log, c("trash", "time_log"), sep = " ") %>% 
  select(home_winter_id, ibid, date_log, time_log)

kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") %>% 
  filter(area == "AK") %>% 
  select(home_winter_id, winter_id, home, home_id)

ak_sums_log <- ak_sums_data_log %>% 
  left_join(kids_linked_ids, by = "home_winter_id") %>% 
  select(home_winter_id, winter_id, home, home_id, date_log, time_log, ibid) %>% 
  arrange(home, date_log)

write_csv(ak_sums_log, "Output/ak_sums_data_log.csv")
write_rds(ak_sums_log, "Output/ak_sums_data_log.rds")

@@

# Bind all AK SUMs data together
ak_sums_raw <- rbind(ak_sums_csv, ak_sums_csv_2, 
                     ak_sums_xlsx, ak_sums_xlsx2) %>% 
  select(-datetime_sums) %>% 
  unite("datetime_sums", c("date_sums", "time_sums"), sep = " ", remove = FALSE) %>% 
  filter(!is.na(date_sums)) %>% 
  arrange(home, datetime_sums) %>% 
  distinct()

# write_rds(ak_sums_raw, "Output/ak_sums_raw.rds")
```

# Bind all SUMs files together
```{r}
wmt_sums_raw <- read_rds("Output/wmt_sums_raw.rds")

nn_sums_raw <- read_rds("Output/nn_sums_raw.rds")

ak_sums_raw <- read_rds("Output/ak_sums_raw.rds")

all_sums_raw <- rbind(wmt_sums_raw, nn_sums_raw, ak_sums_raw)

# write_rds(all_sums_raw, "Output/all_sums_raw.rds")
```

# Check sample start times vs log start times
```{r}
# Load all SUMs data
all_sums_raw <- read_rds("Output/all_sums_raw.rds") %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)

# Load SUMs data logs
wmt_sums_log <- read_rds("Output/wmt_sums_data_log.rds")
nn_sums_log <- read_rds("Output/nn_sums_data_log.rds")
ak_sums_log <- read_rds("Output/ak_sums_data_log.rds")

# Combine SUMs logs from all areas
all_sums_log <- rbind(wmt_sums_log, nn_sums_log, ak_sums_log) %>% 
  mutate(winter_id = as.character(winter_id))

# Combine SUMs logs and data
## Get ready to compare start times from logs and data files
sums_time_comparison <- all_sums_raw %>%
  group_by(area, home) %>% 
  mutate(datetime_initial = first(datetime_sums)) %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums),
         datetime_initial = ymd_hms(datetime_initial),
         winter_id = if_else((datetime_sums - datetime_initial) > 20736000,
                             2, 1), 
         winter_id = as.character(winter_id)) %>% 
  ungroup() %>% 
  left_join(all_sums_log, by = c("home", "winter_id")) %>% 
  group_by(home, winter_id) %>% 
  unite(datetime_log, c("date_log", "time_log"), 
        sep = " ", remove = FALSE) %>% 
  mutate(datetime_log = ymd_hms(datetime_log),
         datetime_initial = first(datetime_sums),
         time_diff_sums = datetime_initial - datetime_log) %>% 
  distinct(time_diff_sums, .keep_all = TRUE) %>%
  ungroup() %>% 
  arrange(home, winter_id)

write_csv(sums_time_comparison, "Output/sums_time_comparison.csv")
write_rds(sums_time_comparison, "Output/sums_time_comparison.rds")
```
